Attributes-based Object Classification and Segmentation

### 2010-05-12 paper.tex
changes

%-------------------------------------------------------------------------------
% abstract
%-------------------------------------------------------------------------------
\begin{abstract}
Attributes have been shown to be effective in assisting object classification and
classification, and they provide means to also incorporate semantic
information available to and provided by humans.  On another front, the
combination of segmentation and object classification has been shown to
improve both segmentation and classification.  We present a method of incorporating
attributes into segmentation using a hierarchical methodology.
That is, using attributes as features, we discover attributes of regions
at all levels of a hierarchical segmentation,
incorporate hierarchical confidences in determining a confidence that
a given region is part of a given object, and segment the image using
these confidences.  We also present encouraging segmentation results
from experiments run using this method.
\end{abstract}

@@

Our work focuses on leveraging attributes to produce accurate segmentations.
The idea is to find image regions with attribute distributions very similar
to some canonical distribution for an object.  Figure \ref{fig:truck} shows
an example.  We define the ``canonical distribution''
to be an attribute distribution representative of that object.  For example, the
canonical distribution for a ``car'' would have some measure of ``shininess'', some
other measure of ``window'' and so forth.  If
regions can be combined together to for larger regions with attribute
distributions matching the object's canonical distribution then one can
declare the combined region an accurate segmentation for that object.

@@

In general, the testing of all combinations of regions to
find objects is intractable.  We use two approximatory algorithms
operating on a segmentation hierarchy to determine regions belonging
to a given class.
One algorithm is additive and operates by combining regions that are likely
to be part of the object.  The other is subtractive, which first predict attributes on
the entire region and then proceeds by observing the effects of
removing different regions during a top-down traversal of the tree.
The benefit of using these methods are that large regions that
clearly don't contribute to an attribute distribution similar to the
canonical can be pruned from
the tree early in the traversal.  This approach avoids the
combinatoric complexity of a full search and, even more, prunes many
regions from the segmentation tree before they even are searched,
bringing the computational complexity, in practice, to roughly
$O(n)$ where $n$ is the number of leaves in the segmentation tree.
The flowchart of our algorithm is shown in Figure
\ref{fig:flowchart}.

This work is done using attributes as the features, but the segmentation
pruning algorithms are actually feature-type agnostic and could be used
with any image feature given a scoring function applicable to that feature.

@@

Attributes are high-level features which can reflect semantic properties of objects. Given the attributes, we can do object detection based on their distribution. We use the same attribute set as \cite{farhadi09}. This set contains three types of attributes, including shape, part and material. All attributes are listed in figure \ref{fig:atts}.

@@

We use a similar algorithm to that used in \cite{farhadi09} to train the attribute models. 
Three different type of features are used: HOG, texture and color descriptors \cite{farhadi09}.  
We use kmeans to quantize the features. The HOG descriptors are quantized into 1000 clusters, 
and the texture and color are quantized to 256 and 128 clusters, respectively. 
All quantized features are concatenated to form a 1384 dimensional feature.
Since all features are local, we represent a segment by using only features inside the segment
to produce a feature histogram.  In other words, rather than segmenting the image and finding features,
we find features and then segment the features.
This is done to avoid introducing any feature artifacts as a result of black space surrounding 
segmented regions in the image.

@@

We use the segmentation approach of \cite{arbelaez09}.  This
produces a hierarchical segmentation by first detecting contours
using the \emph{gPb} detector \cite{maire08}.  This contour detector
is advantageous because it uses brightness, color and texture
gradients at multiple scales.  Once the contours are
detected, regions are discovered using a watershed algorithm after
which an Ultrametric Contour Map (UCM) is constructed, which defines
the region hierarchy.  The UCM is stored as a grayscale image with
contours around the segmented regions.  These contours have intensities
representative of the region's location in the hierarchy.  So to find
the first branch in the tree, threshold the UCM image with a value
$t$ where $t = max{I(x,y)}$ where $I(x,y)$ is the intensity at pixel
location $(x,y)$.  The thresholded image can then be flood-filled into
regions.  These regions correspond to the branches on either side of
the root node in the segmentation tree.  The next level of the tree is
found by thresholding the two new regions of the UCM at each, respective,
next highest intensity and so on.
Figure \ref{fig:ucm} shows an example UCM image and figure \ref{fig:tree}
shows the hierarchy generated from that UCM.

After the construction of segmentation tree, we detect all
the attributes in each node, including the intermediate and
leaf nodes. We use the output of SVM as the confidence (not binary),
so each node can be represented by the confidence distribution of
all attributes.  These distributions are stored and later used in
the additive pruning algorithm.  Unfortunately the subtractive algorithm
requires runtime predictions, and thus cannot use this preprocessing
step.

@@

Our objective is to detect objects and segment them out. Using a segmentation tree
we cast the problem into one of selecting nodes that belong 
to the desired object. We estimate the
probability that a node belongs to object based on the attribute
confidence distribution. For each object, we manually determine a
canonical attribute distribution.  We use the
histogram intersection between the desired distribution and the
estimated confidence distribution in each node as the similarity
measure. The histogram intersection is calculated as follows:

\[ HI(H_{model},H_{pred}) = \frac{|H_{model} \cap H_{pred}|}{|H_{model}|}
\] ,

where $H_{model}$ and $H_{pred}$ represent the desired attribute
distribution and predicted distribution.

@@

The idea of the algorithm is to find the smallest discriminative
regions, either positive or negative. We combine all the smallest
positive regions while remove all the negative ones to form the
final segmentation. We use three threshold for the similarity
measure, $thre\_perf$, $thre\_pos$ and $thre\_neg$ which represent
perfect segment, positive segment and negative segment,
respectively. If the similarity is above $thre\_perf$, which means
the segmentation is perfect for the object, we cut all the nodes
below and return this segment. If the similarity is below
$thre\_neg$, which means the segment does not contain the object, we
cut all the nodes below and return NULL. If the similarity is
between $thre\_perf$ and $thre\_pos$, which means the node
definitely contains the object but the segmentation is not perfect
yet, we keep on checking its children nodes to refine the
segmentation. The final case is that the similarity is between
$thre\_pos$ and $thre\_neg$, which means we are not sure whether
there is an object in the node, so we just keep on searching its
children.

The object classification is based on the output of ADDITIVE\_PRUNE. If the output is not NULL, which means at least one node in the tree has similarity above $thre\_pos$, we think the object is detected in this image.

@@

The algorithm begins at the root of the tree and predicts attributes.
This attribute distribution and its associated similarity $sim_{root}$
to the object's canonical distribution is taken to be
the standard for the image.  That is, any final segmentation should not have
significantly less similarity $sim$ than $sim_{root}$.  The algorithm traverses
the subtractive tree depth-first.  At each node, attributes are predicted for 
a region $r$.  
$r$ is obtained by taking the intersection of the master mask and the inverted
node mask ($r = m_{m} \cap \verb+~+ m_{n}$).  The intuition is this: the master mask is our current best guess at
the region that contains an object.  By taking the intersection, we effectively
subtract the node's region out.
We can then predict on the new region.  If 
there is no significant negative effect on the similarity as
compared to the standard, then we assume that we can safely prune off the node's
region without removing a part of the object..  If the
similarity was significantly affected, then we assume that the region
contains some part of the object and we allow it to remain.  But that 
doesn't mean that sub-regions might not be able to be pruned
out, so the procedure is called recursively, checking for smaller regions
that might be able to be pruned.

@@

In this section, we evaluate our attribute based object classification
and segmentation algorithms. We also analyze the advantages and 
disadvantages of our algorithms.\\

@@

Table \ref{table:aresults} shows the classification results comparing the proposed method with 
the other three baseline methods. LowFea means that we use the low level feature to train and test 
the object classification directly. Att1 means that we use the attribute distribution as the feature,
but we only test it in the whole image without segmentation. Random means random prediction. Add is our
proposed algorithm with additive pruning. The proposed algorithm has the highest precision comparing 
with the other baselines, while the recall is a little lower than LowFea. Figure \ref{fig:roc_class} 
shows the ROC curve for each class and the average.

@@

Figure \ref{fig:roc_seg} shows the segmentation precision of additive pruning 
versus the false positive rate of object classification. The segmentation precision
is in the range of $20\%$ to $50\%$. Some of the segmentation results are shown in 
Figure \ref{fig:car_good_results},\ref{fig:car_bad_results},\ref{fig:aeroplane_good_results},\ref{fig:aeroplane_bad_results},\ref{fig:horse_good_results},\ref{fig:horse_bad_results}.  Again, the subtractive algorithm doesn't produce
analogous results as it performs no classification.

@@

Figures \ref{fig:aeroplane_bad_results} and \ref{fig:horse_bad_results} show results
that either miss part of the object or contain a lot of background.  Extreme cases,
such as the second image in \ref{fig:horse_bad_results} mask out the object of interest
(horse, in this case) and almost nothing else.  These errors are cause by two things:
first, the additive algorithm currently does not use all levels in the segmentation tree,
as it would have made testing take prohibitively long.
Another problem with searching deep in the tree is that when the segments are too small, 
the attribute detection is nearly random, because the small regions 
don't have enough points to represent their properties. Second, the training of attributes based on
bounding boxes is not accurate, because of many background pixels can be included in the bounding boxes. 
This is especially obvious to the horse images, since horses are far from having a convex shape. 
All the segmented images in figure \ref{fig:horse_bad_results} contains grass or trees, especially the 
left one which only contains grass. These images imply that the grass and tree models are actually
included in the horse model.\\

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/add_res/roc_seg.png.eps}
\caption{Result of adjusting the factor $f$ in the subtractive
algorithm. \subref{fig:fa} shows running the algorithm with $f=0.9$
and \subref{fig:fb} shows running the algorithm with $f=1.0$.  Note
that \subref{fig:fb} retains more of the cars, especially the two
closest on the right and left, but also has more false-positive
pixels, such as the ceiling region of the parking garage.}
\label{fig:roc_seg}
\end{figure}

@@

It can be seen that in simple cases both additive and subtractive algorithms
are able to segment out rough outlines of the objects.  In most cases, even
the more difficult ones, at least part of the car was present in the final
segmentation.  This shows that the segmentation algorithms appear to be
reasonable approximations to a combinatoric approach -- that of predicting
on every combination of regions in the image.

Despite these encouraging initial results, the final segmentation region
was often not coherent.
This appears to be largely because the region tree approach is not
constrained by spatial coherence.  That is, two leaf nodes containing image
regions that are spatially next to each other may be very distant in the
region tree.  Because of this, the attribute distribution may change significantly
after considering the first node and before consideration of the second node.
Another reason for this may be that our canonical attribute distribution for
our objects gave equal weights to each attribute.  For example, a wheel was
equally as important as shininess.  This gave unfair disadvantages to
under-represented or ill-trained attributes.  This can be seen in the results
where the body of the car is removed while the wheels or headlights are
retained.

@@

As object classification is generally only as good as the features used, so also it
appears that this approach to segmentation is only as good as the attribute
selection and training.  We found that in images with poor segmentations, the
confidences in prediction of object-related attributes were generally low. Using bounding 
boxes as ground truth can introduce many background features into the attribute models. These 
features can be harmful. The extreme example is shown in the left image of figure \ref{fig:horse_bad_results}, 
which only segments out the background grass. However, even if we have the segmentation 
of objects as the ground truth for training the attributes, we may still not have accurate 
models, since the ground truth is for objects not for attributes. For example, if we train 
the attributes of wheel or metal only based on car images, we may get the car model not 
the attributes model. One way to deal with this problem is to directly label attributes 
ground truth. Then we can get similar accurate attribute models as the object models.

@@

The holy grail of this approach, of course, is to be able to segment any object
with a given canonical attribute distribution.  If this can be achieved, 
then segmentation experiments
could open up to non-experts, who could define an attribute distribution
based on their own knowledge and use a system that has already been trained
on a large number of attributes to segment the object.

>>>>

%-------------------------------------------------------------------------------
% abstract
%-------------------------------------------------------------------------------
\begin{abstract}
Attributes have been shown to be effective in assisting object classification, and they provide means to also incorporate semantic
information available to and provided by humans.  On another front, the
combination of segmentation and object classification has been shown to
improve both segmentation and classification.  We present a method of incorporating
attributes into segmentation using a hierarchical methodology.
That is, using attributes as features, we discover attributes of regions
at all levels of a hierarchical segmentation,
incorporate hierarchical confidences in determining a confidence that
a given region is part of a given object, and segment the image using
these confidences.  We also present encouraging segmentation results
from experiments run using this method.
\end{abstract}

@@

Our work focuses on leveraging attributes to produce accurate segmentations.
The idea is to find image regions with attribute distributions very similar
to some canonical distribution for an object.  Figure \ref{fig:truck} shows
an example.  We define the ``canonical distribution''
to be an attribute distribution representative of that object.  For example, the
canonical distribution for a ``car'' would have some measure of ``shininess'', some
other measure of ``window'' and so forth.  If
regions can be combined together into larger regions with attribute
distributions matching the object's canonical distribution then one can
declare the combined region an accurate segmentation for that object.

@@

In general, the testing of all combinations of regions to
find objects is intractable.  We use two approximatory algorithms
operating on a segmentation hierarchy to determine regions belonging
to a given class.
One algorithm is additive and operates by combining regions that are likely
to be part of the object.  The other is subtractive, which first predicts attributes on
the entire region and then proceeds by observing the effects of
removing different regions during a top-down traversal of the tree.
The benefit of using these methods are that large regions that
clearly don't contribute to an attribute distribution similar to the
canonical can be pruned from
the tree early in the traversal.  This approach avoids the
combinatoric complexity of a full search and, even more, prunes many
regions from the segmentation tree before they even are searched,
bringing the computational complexity, in practice, to roughly
$O(n)$ where $n$ is the number of leaves in the segmentation tree.
The high-level outline of our algorithm is shown in Figure
\ref{fig:flowchart}.

@@

Attributes are high-level features which can reflect semantic properties of objects. Given the attributes, we can perform object classification based on an image's attribute distribution. We use the same attribute set as \cite{farhadi09}. This set contains three types of attributes, including shape, part and material. All attributes are listed in figure \ref{fig:atts}.

@@

We use a similar approach to that used in \cite{farhadi09} to train the attribute models. 
Three different type of features are used: HOG, texture and color descriptors \cite{farhadi09}.  
We use kmeans to quantize the features. The HOG descriptors are quantized into 1000 clusters, 
and the texture and color are quantized to 256 and 128 clusters, respectively. 
All quantized features are concatenated to form a 1384 dimensional feature.
Since all features are local, we represent a segment by using only features inside the segment
to produce a feature histogram.  In other words, rather than segmenting the image and finding features,
we find features and then segment the features.
This is done to avoid introducing any feature artifacts as a result of black space surrounding 
segmented regions in the image.

@@

We use the segmentation approach of \cite{arbelaez09}.  This
produces a hierarchical segmentation by first detecting contours
using the \emph{gPb} detector \cite{maire08}.  This contour detector
is advantageous because it uses brightness, color and texture
gradients at multiple scales.  Once the contours are
detected, regions are discovered using a watershed algorithm after
which an Ultrametric Contour Map (UCM) is constructed, which defines
the region hierarchy.  The UCM is stored as a grayscale image with
contours around the segmented regions.  These contours have intensities
representative of the region's location in the hierarchy.  So to find
the first branch in the tree, we threshold the UCM image with a value
$t$ where $t = max\{I(x,y)\}$ where $I(x,y)$ is the intensity at pixel
location $(x,y)$.  The thresholded image can then be flood-filled into
regions.  These regions correspond to the branches on either side of
the root node in the segmentation tree.  The next level of the tree is
found by thresholding the two new regions of the UCM at each, respective,
next highest intensity and so on.
Figure \ref{fig:ucm} shows an example UCM image and figure \ref{fig:tree}
shows the hierarchy generated from that UCM.

After the construction of segmentation tree, we detect all
the attributes in each node, including the intermediate and
leaf nodes. We use the output of the SVM as the confidence (not binary),
so each node can be represented by the confidence distribution of
all attributes.  These distributions are stored and later used in
the additive pruning algorithm.  Unfortunately the subtractive algorithm
requires runtime predictions, and thus cannot take advantage of this preprocessing
step.

@@

Our objective is to detect objects and segment them out. Using a segmentation tree
we cast the problem into one of selecting nodes that belong 
to the desired object. We estimate the
probability that a node belongs to object based on the attribute
confidence distribution. For each object, we manually determine a
canonical attribute distribution.  We use the
histogram intersection between the canonical distribution and the
estimated confidence distribution in each node as the similarity
measure. The histogram intersection is calculated as follows:

\[ HI(H_{model},H_{pred}) = \frac{|H_{model} \cap H_{pred}|}{|H_{model}|}
\]
where $H_{model}$ and $H_{pred}$ represent the desired attribute
distribution and predicted distribution.

@@

The idea of the algorithm is to find the smallest discriminative
regions, either positive or negative. We combine all 
positive regions while removing all negative ones to form the
final segmentation. We use three thresholds for the similarity
measure, $thre\_perf$, $thre\_pos$ and $thre\_neg$ which represent
perfect segment, positive segment and negative segment,
respectively. If the similarity is above $thre\_perf$, which means
the segmentation is perfect for the object, we cut all descendent nodes
and return this segment. If the similarity is below
$thre\_neg$, which means the segment does not contain the object, we
cut all descendent nodes and return NULL. If the similarity is
between $thre\_perf$ and $thre\_pos$, which means the node
definitely contains the object but the segmentation is not perfect
yet, we check its children nodes to refine the
segmentation. The final case is that the similarity is between
$thre\_pos$ and $thre\_neg$, which means we are not sure whether
there is an object in the node, so we just search its
children.

The object classification is based on the output of ADDITIVE\_PRUNE. If the output is not NULL, which means at least one node in the tree has similarity above $thre\_pos$, then the object is detected in the image.

@@

The algorithm begins at the root of the tree and predicts attributes.
This attribute distribution and its associated similarity $sim_{root}$
to the object's canonical distribution is taken to be
the standard for the image.  That is, any final segmentation should not have
significantly less similarity $sim$ than $sim_{root}$.  The algorithm traverses
the subtractive tree depth-first.  At each node, attributes are predicted for 
a region $r$.  
$r$ is obtained by taking the intersection of the master mask and the inverted
node mask ($r = m_{m} \cap \verb+~+ m_{n}$).  The intuition is this: the master mask is our current best guess at
the region that contains an object.  By taking the intersection, we effectively
subtract the node's region out.
We can then predict on the new region.  If 
there is no significant negative effect on the similarity as
compared to the standard, then we assume that we can safely prune off the node's
region without removing a part of the object.  If the
similarity was significantly affected, then we assume that the region
contains some part of the object and we allow it to remain.  But that 
doesn't mean that sub-regions might not be able to be pruned
out, so the procedure is called recursively, checking for smaller regions
that might be able to be pruned.

@@

In this section, we evaluate our attribute-based object classification
and segmentation algorithms. We also analyze the advantages and 
disadvantages of our algorithms.\\

@@

Table \ref{table:aresults} shows the classification results comparing the proposed method with 
the other three baseline methods. LowFea means that we use the low level features to train and test 
the object classification directly. Att1 means that we use the attribute distribution as the feature,
but we only test it in the whole image without segmentation. Random means random prediction. Add is our
proposed algorithm with additive pruning. The proposed algorithm has the highest precision comparing 
with the other baselines, while the recall is a little lower than LowFea. Figure \ref{fig:roc_class} 
shows the ROC curve for each class and the average.

@@

Figure \ref{fig:roc_seg} shows the segmentation precision of additive pruning 
versus the false positive rate of object classification. The segmentation precision
is in the range of $20\%$ to $50\%$. Some of the segmentation results are shown in 
Figure \ref{fig:car_good_results},\ref{fig:car_bad_results},\ref{fig:aeroplane_good_results},\ref{fig:aeroplane_bad_results},\ref{fig:horse_good_results},\ref{fig:horse_bad_results}.

@@

Figures \ref{fig:aeroplane_bad_results} and \ref{fig:horse_bad_results} show results
that either miss part of the object or contain a lot of background.  Extreme cases,
such as the second image in figure \ref{fig:horse_bad_results} removes the object of interest
(horse, in this case) and almost nothing else.  These errors are caused by two things:
first, the additive algorithm currently does not use all levels in the segmentation tree,
as it would make testing take prohibitively long.
Another problem with searching deep in the tree is that when the segments are too small
the attribute detection is nearly random, because the small regions 
don't have enough features to represent their attributes. Second, the training of attributes based on
bounding boxes is not accurate, because many background pixels can be included in the bounding boxes. 
This is especially obvious to the horse images, since horses are far from having a convex shape. 
All the segmented images in figure \ref{fig:horse_bad_results} contains grass or trees, especially the 
left one which only contains grass. These images imply that the grass and tree models are actually
included in the horse model.\\

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/add_res/roc_seg.png.eps}
\caption{The curve of segmentation precision versus false positive classification rate.}
\label{fig:roc_seg}
\end{figure}

@@

It can be seen that, in simple cases, both additive and subtractive algorithms
are able to segment out rough outlines of the objects.  In most cases, even
the more difficult ones, at least part of the object was present in the final
segmentation.  This shows that the segmentation algorithms appear to be
reasonable approximations to a combinatoric approach -- that of predicting
on every combination of regions in the image.

Despite these encouraging initial results, the final segmentation region
was often not coherent.
This appears to be largely because the region tree approach is not
constrained by spatial coherence.  That is, two leaf nodes containing image
regions that are spatially next to each other may be very distant in the
region tree.  Because of this, the attribute distribution may change significantly
after considering the first node and before consideration of the second node.
Another reason for this may be that our canonical attribute distribution for
our objects gave equal weights to each attribute.  For example, a wheel was
equally as important as shininess.  This gave unfair disadvantages to
under-represented or ill-trained attributes.

@@

As object classification is generally only as good as the features used, so also it
appears that this approach to segmentation is only as good as the attribute
selection and training.  We found that in images with poor segmentations, the
confidences in prediction of object-related attributes were generally low. Using bounding 
boxes as ground truth can introduce many background features into the attribute models. The extreme example is shown in the left image of figure \ref{fig:horse_bad_results}, 
which only segments out the background grass. However, even if we have the segmentation 
of objects as the ground truth for training the attributes, we may still not have accurate 
models, since the ground truth is for objects, not for attributes. For example, if we train 
the attributes of wheel or metal only based on car images, we may get the car model rather than 
the attributes model. One way to deal with this problem is to directly label attributes' 
ground truth. Then we can get similar accurate attribute models as the object models, although 
it would be a tremendously time-consuming task to manually segment all attributes' regions.

@@

The holy grail of our approach, of course, is to be able to segment any object
with a given canonical attribute distribution.  If this can be achieved, 
then segmentation experiments
could open up to non-experts, who could define an attribute distribution
based on their own knowledge and use a system that has already been trained
on a large number of attributes to segment the object.
